### 2018-05-23 doc/submission/reviewer-response.md

    *We thank the reviewer for noticing that omission. The r values have been
    included in the text.*

# Reviewer 2

1.	It is a frustrating that this is seemingly the third in a series of papers
that is looking at patterns in lipid fractions and association with metabolic
outcomes. Thus, the TG, NEFA and phospholipid fractions all seem to be a 'least
publishable unit'. The authors must integrate the knowledge gained from the
previous analysis of NEFA and phospholipids and discuss some of the seemingly
contradictory findings from the previous studies. If the data were examined,
would the DNL story 'hold up'?

    *We thank the reviewer for these comments. We agree that ideally the results
    of all the pools would have been included together, but given the number of
    fatty acids, the longitudinal visits, the depth and volume of the analyses,
    and the complex biology involved, it simply wasn't possible to contain
    everything in a single journal paper. We have included a brief discussion of
    the results from the other fractions in the Discussion section. The 
    complete and integrated discussion and presentation of the results is found
    in the first author's (LWJ) PhD thesis (http://hdl.handle.net/1807/80893). As to a combined analysis,
    we have presented these results at the European Diabetes Epidemiology Group
    Meeting in April, 2018 (https://doi.org/10.6084/m9.figshare.6159293), and the proposed DNL mechanism
    remained clearly associated and had a strong contribution to associations
    with diabetes pathogenesis compared to other fatty acids/compositions.
    A short aside, these currently unpublished findings do not invalidate or 
    undermine the previous work and the current manuscript. The combined analysis uses a powerful
    statistical technique (PLS), but it has a number of major limitations,
    not least of which is it can only handle cross-sectional data and can not
    adjust for potential confounders. The present analysis in this manuscript uses PLS in addition
    to the powerful longitudinal technique GEE.*

### 2018-05-28 doc/manuscript.Rmd
todo -> done, add NEFA and other fractions text

<!--
{{ Not sure about this... Include this? }}
As with our previous analysis of the PL fatty acid fraction {{cite}}, we found
that cis-vaccenic acid strongly predicted higher insulin sensitivity and
beta-cell function. Likewise, in the TGFA fraction we found a similar
association with cis-vaccenic acid. {{we can discuss to include more}}

{{ comment on comparing models with and without NEFA }} While
NEFA may be a mediating factor, we found that adjustment for it did not
{{confirm}} attenuate this association since TGFA composition is also dictated
by other sources for DNL (e.g. recent meal, DNL from glucose stores, etc.).

While the NEFA fraction is much smaller than the TG fraction, even though it
contributes substantially to the TG fraction, the VLDL-TG is not cleared
(lipolyzed) as much as chylomicron (which also contains a large amount of TG)
and so has a higher half-life than NEFA, remaining in circulation longer {{hence
why there is bigger TG pool vs NEFA}} {{cite?}}.

@@

{{ Concentration data is not }}

>>>>

Given the close biological relationship between circulating NEFA and TAG, NEFA 
may act as confounding factor and was thus adjusted for. In a previous published
analysis of the NEFA fraction [@Johnston2017b], we found that higher total NEFA,
but not the specific composition, associated with lower beta-cell function. This
is in contrast to the TGFA findings that the specific composition does
differentially associate with insulin sensitivity and beta-cell function,
adjusting for total NEFA. There was no difference in results in models that did
not include NEFA as a confounder (data not shown). This difference in results
between NEFA and TGFA suggests that TGFA may independently and strongly influence the 
pathophysiology of T2D, when compared to other lipid fraction compositions,
including the phospholipid and cholesteryl ester fractions [@Johnston2016a]. This
may be because TAG are biologically destined for uptake by non-hepatic tissue as
they are found mainly in VLDL, at least during fasting. This is in contrast to
NEFA that are mostly uptaken by the liver and used in TAG production [@Nielsen2012a].

<!--
{{ comment on comparing models with and without NEFA }} While NEFA may be a
mediating factor, we found that adjustment for it did not attenuate this
association since TGFA composition is also dictated by other sources for DNL
(e.g. recent meal, DNL from glucose stores, fatty acid recycling, etc.).

@@
